\sysfull

### 2017-08-09 paper/coalescing.tex
wrote the coalescing section.  we need the figure

\sys can coalesce an arbitrary number consecutive tasks.  As more tasks
coalesce, their collective commit overhead amortizes better, potentially
improving performance.  As more tasks coalesce, the amount of potentially {\em
wasted work} also increases. If power fails during a long sequence of coalesced
tasks the lack of an intervening commit requires execution to restart from the
{\em first} task in the sequence, failing to preserve the progress of any of
the tasks in the long sequence.

\subsection{Task Coalescing Algorithm}

\begin{algorithm}[t]
	\caption{\sys task coalescing mechanism}
	\label{algo:coalescing}
	\scriptsize
	\begin{algorithmic}[1]
		\State $\text{p}$  \Comment{Indicates persistent variables, initialized \textbf{only} during device programming}
		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
		\State $\text{VT} \subset \{~\text{RT's}\}$  \Comment{Virtual task}
		\State $\text{RTC}$  \Comment{RT counter}
		\State $\text{VTC}$  \Comment{VT counter}
		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
		\State $\text{TVT}$ \Comment{Temporary VT}
		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
		\vspace{0.1cm}

		\State $\text{p RTC} = x $ 
		\State $\text{p VT}_{\max} = y$
		\State $\text{p VT} = \text{RT}$ 
		\State $\text{p TVT} = \text{VT} $ 
		\State $\text{VTC} = 0 $ 
		\vspace{0.1cm}

		\Function{scheduler}{\null}

			\State $\text{VTC} = \text{RTC/2} $  \label{algo:coalescing:executionHistory1}
			\State $\text{RTC} = 0 $
			\State $\text{RT} = \text{VT}$ \Comment{Recover virtual state} \label{algo:coalescing:virtualProgressing1}
			\If{commiting}
				\State $\text{RT}=\text{TVT}$
				\State $\texttt{goto} \ \ \ref{commitStage}$ \label{algo:coalescing:firmTransition1}
			\EndIf
			\vspace{0.1cm}

			\While {True}
				\While{$\text{VTC}--$}
					\Function{execute}{$\text{RT}$}
					\EndFunction
					\If{$\text{RTC} < \text{VT}_{\max}$}
						\State $\text{RTC}++$  \label{algo:coalescing:realTaskCounter}
					\EndIf
					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{Virtual progressing}
				\EndWhile        \label{algo:coalescing:virtualProgressing2}

				\State \Comment{Virtual task is finished}
				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temporary buffer}
				\State $\text{TVT} = \text{RT}$ 
				\State $\text{commiting} = \text{True}$ \label{algo:coalescing:firmTransition2}
				\State 
				\State $\text{VT} = \text{TVT}$ \label{commitStage}
				\State $\text{pTempBuf} \rightarrow \text{FRAM}$ 
				\State $\text{VTC} = \text{RTC/2}$ 		\Comment{Set VT size} \label{algo:coalescing:executionHistory2}
				\State $\text{commiting} = \text{False}$

			\EndWhile

		\EndFunction

	\end{algorithmic}
\end{algorithm}

\sys's coalescing algorithm uses a single point back in the \sys's execution
history to estimate the appropriate virtual task size. However, it takes a
conservative approach by making the virtual task size equal to the half of the
number of successfully executed real tasks before the last power interrupt (see
lines~\ref{algo:coalescing:executionHistory1},
\ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}). We
advocate for the algorithm conservative approach by the following two reasons:
(i) the power interrupts are not uniformly distributed over the tasks; and (ii)
the re-execution penalty, in case of a power interrupt, grows linearly with the
size of a virtual task. Furthermore, the algorithm considers the charging rate
by allowing the real tasks counter (RTC), which represents the execution
history between the last two power interrupts, to increase from one up to
multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of
Algorithm~\ref{algo:coalescing}). As a consequence, the virtual task size
changes rapidly (since it is half of the execution history) in response to the
change in the charging rate. However, the algorithm sets a limit to the size of
the virtual task for these two reasons: (i) having an infinity virtual task
size, results in a certain computation progress loss; and (ii) the benefit of
task merging is governed by the diminishing return principle (which will be
demonstrated experimentally in Section~\ref{sec:results_coalescing}).

The algorithm protects itself from power interrupts by going into three stages: (i) virtual progressing, where the operations are done on the volatile memory (see lines~\ref{algo:coalescing:virtualProgressing1}--\ref{algo:coalescing:virtualProgressing2} of Algorithm~\ref{algo:coalescing} ); (ii) transition stage, where the volatile state is moved to a temporary persistent buffer. Until the end of the second stage if a power is interrupted, then all the virtual progress will be cancelled and the virtual task will re-executed and re-initialized from a consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of Algorithm~\ref{algo:coalescing}); and (iii) persistent commit stage, upon entering this stage the algorithm will make a firm transition (see lines~\ref{algo:coalescing:firmTransition1}, \ref{algo:coalescing:firmTransition2} of Algorithm~\ref{algo:coalescing}) and it will not go back to the other stages unless all the data is committed from the persistent buffer to non-volatile memory. Since the data is being moved in one direction and from a persistent location, power interrupt is tolerable at this stage.

\subsection{Power Interrupt Immune Scheduler}

\input{relativeJumpAlgo.tex} %Task jumping algorithm

It utilizes a persistent circular buffer (i.e. persistent linked list) to keep the state of a program across power failures. \sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e. (un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop. \todo{Expand this section}{Amjad}

%\begin{algorithm}
%	\caption{Opportunistic virtual Task size}
%	\label{algo:fixVirtTask}
%	\scriptsize
%	%\small
%	\begin{algorithmic}[1]
%		\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
%		\State VTS : VT size
%		\vspace{0.1cm}
%		
%		\While {$True$}
%		\State $VT \leftarrow VT_{next}$
%		\vspace{0.1cm}
%		\While {execute $VT$} 
%		\If { $\text{power failed twice}$ }				
%		\State $VTS--$  
%		\EndIf
%		\EndWhile
%		
%		\vspace{0.1cm}
%		\If {$ \text{All tasks executed}$}
%		\State $VTS++$
%		\EndIf
%		\EndWhile
%	\end{algorithmic}
%\end{algorithm}

>>>>

\sys can coalesce an arbitrary number consecutive tasks.  As more tasks
coalesce, their collective commit overhead amortizes better, potentially
improving performance.  As more tasks coalesce, the amount of potentially {\em
wasted work} also increases. If power fails during a long sequence of coalesced
tasks the lack of an intervening commit requires execution to restart from the
{\em first} task in the sequence, failing to preserve the progress of any of
the tasks in the long sequence. \sys moderates the risk of coalescing and
capitalizes on its benefit by dynamically, adaptively determining a coalescing
factor that combines many tasks without suffering an excess of wasted work.

The key challenge to coalescing tasks is determining how many tasks to coalesce
before committing.  During execution, an executing task is either a {\em
coalesced} task, which does not commit its state before transitioning or a {\em
committing} task, which does commit its state before transitioning. \sys
monitors a program's execution to adaptively determine $C$, the number of tasks
to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
followed by a committing task.

\sys determines $C$ for a given operating based on the number of tasks
completed during the last operational period.  \sys begins execution in a
coalesced task with an infinite $C$ value.  With a continuous power supply, the
system executes from the start with all tasks as coalesced tasks entirely
avoiding the cost of committing state.  On intermittent power, the system
executes a sequence of tasks that is interrupted by a power failure.  \sys
explicitly tracks $S$, the number of tasks completed during the previous
operating period.  \sys assigns $C$ to be equal to half of the value of $S$.
If more tasks in total are able to complete in an operating period, \sys
attempts to coalesce more tasks during the next operating period.  If fewer
tasks complete, \sys attempts to coalesce fewer tasks during the next operating
period. $S$ includes the count of both coalesced and committing tasks
completed.  If no committing task completes (i.e., execution does not make
progress), \sys updates $S$ and attempts to coalesce fewer tasks during the
next operating period.

Figure~\ref{fig:coalescing} illustrates how \sys coalesces tasks.
Figure~\ref{fig:coalescing}(a) shows a sequence of six tasks executing in \sys
with no coalescing, incurring the cost of commit at each task transition.
Figure~\ref{fig:coalescing}(b) shows the same sequence of tasks executing as
coalesced groups of three tasks. Figure~\ref{fig:coalescing}(c) shows how \sys
adapts the value of $C$ based on $S$.  \sys executes 12 tasks in the first
operating period, assigning $S$ to six.  \sys completes only five of the six
coalesced tasks in the next operating period and then power fails. \sys sets
$S$ to two in the next operating period, coalescing pairs of tasks and
successfully committing two pairs before power fails. Execution resumes at the
start of the third pair of tasks with a coalescing factor of two.  If a larger
number of tasks successfully complete in the next operating period (i.e., $S$
is larger), $C$ correspondingly increases, allowing more aggressive coalescing.

%
%However, it takes a conservative approach by making the virtual task size equal
%to the half of the number of successfully executed real tasks before the last
%power interrupt (see lines~\ref{algo:coalescing:executionHistory1},
%\ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}). 
%
%We
%advocate for the algorithm conservative approach by the following two reasons:
%(i) the power interrupts are not uniformly distributed over the tasks; and (ii)
%the re-execution penalty, in case of a power interrupt, grows linearly with the
%size of a virtual task. 
%
%Furthermore, the algorithm considers the charging rate
%by allowing the real tasks counter (RTC), which represents the execution
%history between the last two power interrupts, to increase from one up to
%multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of
%Algorithm~\ref{algo:coalescing}). 
%
%As a consequence, the virtual task size
%changes rapidly (since it is half of the execution history) in response to the
%change in the charging rate. 
%
%However, the algorithm sets a limit to the size of
%the virtual task for these two reasons: (i) having an infinity virtual task
%size, results in a certain computation progress loss; and (ii) the benefit of
%task merging is governed by the diminishing return principle (which will be
%demonstrated experimentally in Section~\ref{sec:results_coalescing}).
%
%The algorithm protects itself from power interrupts by going into three stages:
%(i) virtual progressing, where the operations are done on the volatile memory
%(see
%lines~\ref{algo:coalescing:virtualProgressing1}--\ref{algo:coalescing:virtualProgressing2}
%of Algorithm~\ref{algo:coalescing} ); 
%(ii) transition stage, where the volatile
%state is moved to a temporary persistent buffer. 
%
%Until the end of the second
%stage if a power is interrupted, then all the virtual progress will be
%cancelled and the virtual task will re-executed and re-initialized from a
%consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of
%Algorithm~\ref{algo:coalescing}); 
%
%and (iii) persistent commit stage, upon
%entering this stage the algorithm will make a firm transition (see
%lines~\ref{algo:coalescing:firmTransition1},
%\ref{algo:coalescing:firmTransition2} of Algorithm~\ref{algo:coalescing}) and
%it will not go back to the other stages unless all the data is committed from
%the persistent buffer to non-volatile memory. 
%
%Since the data is being moved in one direction and from a persistent location,
%power interrupt is tolerable at this stage.

%\begin{algorithm}[t]
%	\caption{\sys task coalescing mechanism}
%	\label{algo:coalescing}
%	\scriptsize
%	\begin{algorithmic}[1]
%		\State $\text{p}$  \Comment{Indicates persistent variables, initialized \textbf{only} during device programming}
%		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
%		\State $\text{VT} \subset \{~\text{RT's}\}$  \Comment{Virtual task}
%		\State $\text{RTC}$  \Comment{RT counter}
%		\State $\text{VTC}$  \Comment{VT counter}
%		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
%		\State $\text{TVT}$ \Comment{Temporary VT}
%		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
%		\vspace{0.1cm}
%		
%		\State $\text{p RTC} = x $ 
%		\State $\text{p VT}_{\max} = y$
%		\State $\text{p VT} = \text{RT}$ 
%		\State $\text{p TVT} = \text{VT} $ 
%		\State $\text{VTC} = 0 $ 
%		\vspace{0.1cm}
%
%		\Function{scheduler}{\null}
%
%			\State $\text{VTC} = \text{RTC/2} $  \label{algo:coalescing:executionHistory1}
%			\State $\text{RTC} = 0 $
%			\State $\text{RT} = \text{VT}$ \Comment{Recover virtual state} \label{algo:coalescing:virtualProgressing1}
%			\If{commiting}
%				\State $\text{RT}=\text{TVT}$
%				\State $\texttt{goto} \ \ \ref{commitStage}$ \label{algo:coalescing:firmTransition1}
%			\EndIf
%			\vspace{0.1cm}
%
%			\While {True}
%				\While{$\text{VTC}--$}
%					\Function{execute}{$\text{RT}$}
%					\EndFunction
%					\If{$\text{RTC} < \text{VT}_{\max}$}
%						\State $\text{RTC}++$  \label{algo:coalescing:realTaskCounter}
%					\EndIf
%					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{Virtual progressing}
%				\EndWhile        \label{algo:coalescing:virtualProgressing2}
%
%				\State \Comment{Virtual task is finished}
%				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temporary buffer}
%				\State $\text{TVT} = \text{RT}$ 
%				\State $\text{commiting} = \text{True}$ \label{algo:coalescing:firmTransition2}
%				\State 
%				\State $\text{VT} = \text{TVT}$ \label{commitStage}
%				\State $\text{pTempBuf} \rightarrow \text{FRAM}$ 
%				\State $\text{VTC} = \text{RTC/2}$ 		\Comment{Set VT size} \label{algo:coalescing:executionHistory2}
%				\State $\text{commiting} = \text{False}$ 
%
%			\EndWhile
%
%		\EndFunction
%			
%	\end{algorithmic}
%\end{algorithm}

%\subsection{Power Interrupt Immune Scheduler}
%
%\input{relativeJumpAlgo.tex} %Task jumping algorithm
%
%It utilizes a persistent circular buffer (i.e. persistent linked list) to keep the state of a program across power failures. \sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e. (un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop. \todo{Expand this section}{Amjad}
%
%\begin{algorithm}
%	\caption{Opportunistic virtual Task size}
%	\label{algo:fixVirtTask}
%	\scriptsize
%	%\small
%	\begin{algorithmic}[1]
%		\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
%		\State VTS : VT size
%		\vspace{0.1cm}
%		
%		\While {$True$}
%		\State $VT \leftarrow VT_{next}$
%		\vspace{0.1cm}
%		\While {execute $VT$} 
%		\If { $\text{power failed twice}$ }				
%		\State $VTS--$  
%		\EndIf
%		\EndWhile
%		
%		\vspace{0.1cm}
%		\If {$ \text{All tasks executed}$}
%		\State $VTS++$
%		\EndIf
%		\EndWhile
%	\end{algorithmic}
%\end{algorithm}
